---
title: "p8105_hw5_fx2206"
author: "Fei Xue"
date: "2025-11-13"
output: github_document
---
```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5)
library(tidyverse)
library(broom)
library(purrr)
library(readr)
library(stringr)

set.seed(8105)
theme_set(theme_bw(base_size = 12))
```

**Problem 1**
```{r}
has_dup_bday <- function(n, days_in_year = 365) {
bdays <- sample.int(days_in_year, size = n, replace = TRUE)
any(duplicated(bdays))
}
```

```{r}
n_grid <- 2:50
n_sims <- 10000

p1_res <- map_df(n_grid, ~{tibble(n = .x, dup = replicate(n_sims, has_dup_bday(.x)))}) |>
group_by(n) |>
summarise(prob_shared = mean(dup), .groups = "drop")

p1_res |> head()
```

```{r}
ggplot(p1_res, aes(n, prob_shared)) +
geom_line() +
geom_point() +
scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
labs(
x = "Group size (n)",
y = "Probability of ≥1 shared birthday",
title = "Birthday paradox simulation (10,000 reps per n)"
)
```
The plot shows how the probability that at least two people in a group share the same birthday increases as group size n grows.
When n is small, the probability of a shared birthday is very low.
However, it rises rapidly after n ≈ 15, crossing 50% around n = 23.
By n ≈ 40, the probability exceeds 90%, and by n = 50, it is nearly 100%.

**Problem 2**
```{r}
set.seed(8105)
n <- 30
sigma <- 5
mus <- 0:6
B <- 5000

simulate_once <- function(mu, n = 30, sigma = 5) {
  x <- rnorm(n, mean = mu, sd = sigma)
  tt <- t.test(x, mu = 0)
  tidy(tt) |>
    mutate(mu_true = mu) |>
    select(mu_true, estimate, p.value) |>
    rename(xbar = estimate, p = p.value)
}

p2_res <- map_dfr(mus, function(mu) {
  reps <- replicate(B, simulate_once(mu, n, sigma), simplify = FALSE)
  bind_rows(reps)
})

p2_res %>% count(mu_true)

```

calculate and draw power graph
```{r}
p2_power <-
  p2_res %>%
  group_by(mu_true) %>%
  summarise(power = mean(p < 0.05), .groups = "drop")

ggplot(p2_power, aes(x = mu_true, y = power)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "True mean (μ)",
    y = "Power (Pr(reject H₀))",
    title = "Power vs. true mean"
  )
```
Power increases monotonically with the effect size (the farther μ is from 0, the more likely we reject the false null).

average simulate value
```{r}
p2_means_overall <-
  p2_res %>%
  group_by(mu_true) %>%
  summarise(avg_xbar = mean(xbar), .groups = "drop")

p2_means_reject <-
  p2_res %>%
  filter(p < 0.05) %>%
  group_by(mu_true) %>%
  summarise(avg_xbar_reject = mean(xbar), .groups = "drop")

p2_means <- left_join(p2_means_overall, p2_means_reject, by = "mu_true")

ggplot(p2_means, aes(mu_true, avg_xbar)) +
  geom_point() +
  geom_line() +
  geom_point(aes(y = avg_xbar_reject), shape = 17, size = 2) +
  geom_line(aes(y = avg_xbar_reject), linetype = 2) +
  labs(
    x = "True mean (μ)",
    y = "Average estimate",
    title = "Average x̄ overall (solid) vs. conditional on rejection (dashed)"
  )
```
The overall sample average of μ across all simulations is approximately equal to the true value of μ, as shown by the solid line lying nearly on the 45-degree reference. This indicates that the estimator μ is unbiased for all values of μ.
However, the average μ conditional on rejecting the null hypothesis (the dashed line) is larger than the true μ when the true mean is small (e.g., μ = 0–3). This occurs because we only keep samples that happen to produce large estimated means that pass the significance threshold (p < 0.05). In other words, this is a form of selection bias — significant results tend to overestimate the true effect size.
When the true μ is large (e.g., ≥ 4), power is near 100%, so nearly all samples reject the null, and the conditional average aligns again with the true mean.

**Problem 3**
```{r}
homs_raw <- read_csv("homicide-data.csv", show_col_types = FALSE)
glimpse(homs_raw)
```
The Washington Post homicide dataset contains records of individual homicide cases from 50 large U.S. cities, including information such as victim demographics, location, and case disposition (e.g., “Closed by arrest,” “Closed without arrest,” “Open/No arrest”).


(1)Describe and create city_state, then summarize totals, unsolved
```{r}
homs <-
  homs_raw |>
  mutate(
    city_state = str_c(city, ", ", state),
    unsolved = disposition %in% c("Closed without arrest", "Open/No arrest")
  )

city_summ <-
  homs |>
  group_by(city_state) |>
  summarise(
    total = n(),
    unsolved = sum(unsolved, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(total))

city_summ |> slice_head(n = 10)

```


(2)Baltimore, MD – prop.test and tidy extraction
```{r}
balt <- city_summ |> filter(city_state == "Baltimore, MD")
balt_pt <- prop.test(x = balt$unsolved, n = balt$total)

balt_tidy <- tidy(balt_pt) |>
  transmute(
    city_state = "Baltimore, MD",
    prop_unsolved = estimate,
    conf_low = conf.low,
    conf_high = conf.high
  )

balt_tidy

```

(3)Run prop.test for every city and extract estimates
```{r}
city_props <-
  city_summ |>
  mutate(
    pt = map2(unsolved, total, ~prop.test(.x, .y)),
    pt_tidy = map(pt, ~tidy(.x))
  ) |>
  select(city_state, pt_tidy) |>
  unnest(pt_tidy) |>
  transmute(
    city_state,
    prop_unsolved = estimate,
    conf_low = conf.low,
    conf_high = conf.high
  )
city_props |> arrange(desc(prop_unsolved)) |> slice_head(n = 5)
```

(4)Plot estimates + CIs per city (ordered by estimate)
```{r}
city_props_plot <-
  city_props |>
  mutate(city_state = fct_reorder(city_state, prop_unsolved))

ggplot(city_props_plot, aes(x = prop_unsolved, y = city_state)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low, xmax = conf_high), height = 0) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Proportion unsolved (95% CI)",
    y = NULL,
    title = "Unsolved homicide proportions by city (prop.test, 95% CI)"
  ) +
  theme(axis.text.y = element_text(size = 9))
```

